---
title: "Mass Spectrometry Data on ExperimentHub"
author:
- name: Laurent Gatto
package: MsDataHub
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Mass Spectrometry Data on ExperimentHub}
  %\VignetteEngine{knitr::rmarkdown}
  %%\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Metabolomics}
  %\VignetteEncoding{UTF-8}
bibiliography: references.bib
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

# Available data

## TripleTOF SWATH

- Type: Raw MS data
- Files: `PestMix1_DDA.mzML` and `PestMix1_SWATH.mzML`
- More details: `?TripleTOF_SWATH"

Load with

```{r, eval = FALSE}
PestMix1_DDA.mzML()
PestMix1_SWATH.mzML()
```

## sciex

- Type: Raw MS data
- Files: `20171016_POOL_POS_1_105-134.mzML` and `20171016_POOL_POS_3_105-134.mzML`
- More details: `?sciex"

Load with

```{r, eval = FALSE}
20171016_POOL_POS_1_105-134.mzML()
20171016_POOL_POS_3_105-134.mzML()
```

## PXD000001

- Type: Raw MS data an MS2 identification
- Files:
  `"TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01-20141210.mzML.gz`
  and
  `TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01-20141210.mzid`

Load with

```{r, eval = FALSE}
TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01-20141210.mzML.gz()
TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01-20141210.mzid()
```

# Session information {-}

```{r sessioninfo, echo=FALSE}
sessionInfo()
```

# References {-}
