---
title: "Mass Spectrometry Data on ExperimentHub"
author:
- name: Laurent Gatto
package: MsDataHub
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Mass Spectrometry Data on ExperimentHub}
  %\VignetteEngine{knitr::rmarkdown}
  %%\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Metabolomics}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

# Introduction

The `MsDataHub` package provides example mass spectrometry data,
peptide spectrum matches or quantitative data from proteomics and
metabolomics experiments. The data are served through the
`ExperimentHub` infrastructure, which allows download them only ones
and cache them for further use. Currently available data are summarised
in the table below and details in the next section.

```{r data}
library("MsDataHub")
DT::datatable(MsDataHub())
```

# Installation

To install the package:

```{r install1, eval = FALSE}
if (!require("BiocManager"))
    install.packages("BiocManager")

BiocManager::install("MsDataHub")
```

To install the package from GitHub:

```{r install2, eval = FALSE}
if (!require("BiocManager"))
    install.packages("BiocManager")

BiocManager::install("RforMassSpectrometry/MsDataHub")
```

# Available data

## TripleTOF

- Type: Raw MS data
- Files: `PestMix1_DDA.mzML` and `PestMix1_SWATH.mzML`
- More details: `?TripleTOF`

Load with

```{r, eval = FALSE}
PestMix1_DDA.mzML()
PestMix1_SWATH.mzML()
```

## sciex

- Type: Raw MS data
- Files: `20171016_POOL_POS_1_105-134.mzML` and `20171016_POOL_POS_3_105-134.mzML`
- More details: `?sciex`

Load with

```{r, eval = FALSE}
20171016_POOL_POS_1_105-134.mzML()
20171016_POOL_POS_3_105-134.mzML()
```

## PXD000001

- Type: Raw MS data and peptide spectrum matches
- Files:
  `TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01-20141210.mzML.gz`
  and
  `TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01-20141210.mzid`
- More details: `?PDX000001`

Load with

```{r, eval = FALSE}
TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01-20141210.mzML.gz()
TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01-20141210.mzid()
```

## CPTAC

- Type: tab-delimited quantitative proteomics data tables (as produced
  by MaxQuant)
- Files: `cptac_a_b_c_peptides.txt`, `cptac_a_b_peptides.txt` and
  `cptac_peptides.txt`
- More details: `?cptac`

Load with

```{r, eval = FALSE}
cptac_a_b_c_peptides.txt()
cptac_a_b_peptides.txt()
cptac_peptides.txt()
```

## FAAH KO

- Type: Raw MS data, in netCDF format.
- File: `ko15.CDF`
- More details: `?cdf`

Load with

```{r, eval = FALSE}
ko15.CDF()
```

# Adding data to `MsDataHub`

1. If you would like additional dataset to `MsDataHub`, start by
   opening an
   [issue](https://github.com/rformassspectrometry/MsDataHub/issues)
   in the package's GitHub repository and describe the new data. In
   particular, provide information about it's provenance, its use, its
   format(s) and acknowledge that the data may be shared freely with
   the community without any restrictions. You may provide an open
   licence specifying the terms it can be re-used, typically a
   CC-BY-SA license.
2. By contribution to the package, you acknowledge that you will
   comply to the R for Mass Spectrometry project [code of
   conduct](https://rformassspectrometry.github.io/RforMassSpectrometry/articles/RforMassSpectrometry.html#code-of-conduct).
3. A maintainer of the package will reply to your issue, confirming
   that the data can be added.
4. At this point, if you are familiar with the development of
   `ExperimentHub` packages and GitHub *pull requests*, you may
   directly send one that adds your data to the package. Make sure (1)
   add appropriate references in the manual page and (2) to add
   yourself as a contributor of the package in the DESCRIPTION file.
5. Alternatively, a maintainer will add the dataset to the package and
   may require your input to make sure the documentation file is
   complete.

# Session information {-}

```{r sessioninfo, echo=FALSE}
sessionInfo()
```
